Automated Lensing Learner - I: An automated Strong Lensing Identification Pipeline

### 2017-03-22 ms.tex
Improved some of the wording about ease of scalability and use on a personal laptop for logistic regression + HOG.

%-------------------------------------------------%
\begin{abstract} 
Gravitational lensing offers a direct probe of the underlying mass
distribution of lensing systems, a window to the high redshift
universe, and a geometric probe of cosmological models.  The advent of
large scale surveys such as the Large Synoptic Sky Telescope and
Euclid has prompted a need for automatic and efficient identification
of strong lensing systems.  We present {\em (ALL) Automated Lensing
  Learner}, a strong lensing identification pipeline that will be
publicly released as open source software.  In this first application,
we employ a fast feature extraction method, Histogram of Oriented
Gradients (HOG), to capture edge patterns that are characteristic of
strong gravitational arcs in galaxy-galaxy lensing.  We use logistic
regression to train a supervised classifier model on the HOG of HST-
and LSST-like images.  We use an area under the curve (AUC) of a
Receiver Operating Characteristic curve to assess model performance;
$AUC=1.0$ is an ideal trained classifier, and $AUC=0.5$ does not
provide any classification information.  Our best performing models on
a training set of 10,000 lens containing images and 10,000 non-lens
containing images exhibit an AUC of 0.975 for an HST-like sample.
However, for one orbit of LSST, our model only reaches an AUC of
0.625.  For 10-year mock LSST observations, the AUC improves to 0.809.
Model performance appears to continually improve with the size of the
training set.  Models trained on fewer images perform better in
absence of the lens galaxy, which can decrease the false positive rate
by blending with light from a lensed galaxy image near the central
lensing galaxy.  However, with larger training data sets, information
from the lensing galaxy improves model performance.  \wording{Our
  results show efficient and effective methods for automated strong
  lensing identification with scalable linear methods that are easy to
  parallelize with existing open source tools and are reproducible on
  personal laptop computers.}
\end{abstract}
%-------------------------------------------------%

@@

%  Relevance of our work
\wording{While there has been a recent surge in the use of neural
  networks applied to astronomy image classification problems, these
  implementations are not readily scalable, nor are the necessary
  computational resources and expertise easily accessible to the
  entire scientific community.  We present a supervised classification
  pipeline built with open source tools, tested on mock HST and LSST
  data, which will also be made publicly available.  Our goal is to
  centralize an open source pipeline that is general enough for the
  user to select or add appropriate image processing techniques to
  augment the features of the lensed image, train and test data with
  different machine learning techniques, and to select the sequence of
  methods that provides optimal model performance.  Results are
  reproducible on personal computers, as both the pipeline and the
  data will be publicly distributed.}

@@

% Benefits of our method
\wording{We emphasize that classical linear classifiers, such as
  logistic regression, is scalable and relatively easy to parallelize
  with open source tools such as {\em
    Spark}\footnote{http://spark.apache.org/}.  Initial checks
  indicated that HOG features are well separated enough for a linear
  classifier algorithm.  Finally, our methods enable reproducible
  results that do not require supercomputing resources.

% Impact of work
These results indicate that effective image classification of
galaxy-galaxy strong lensing systems do not require large scale
computing resources or expertise.}

>>>>

%-------------------------------------------------%
\begin{abstract} 
Gravitational lensing offers a direct probe of the underlying mass
distribution of lensing systems, a window to the high redshift
universe, and a geometric probe of cosmological models.  The advent of
large scale surveys such as the Large Synoptic Sky Telescope and
Euclid has prompted a need for automatic and efficient identification
of strong lensing systems.  We present {\em (ALL) Automated Lensing
  Learner}, a strong lensing identification pipeline that will be
publicly released as open source software.  In this first application,
we employ a fast feature extraction method, Histogram of Oriented
Gradients (HOG), to capture edge patterns that are characteristic of
strong gravitational arcs in galaxy-galaxy lensing.  We use logistic
regression to train a supervised classifier model on the HOG of HST-
and LSST-like images.  We use an area under the curve (AUC) of a
Receiver Operating Characteristic curve to assess model performance;
$AUC=1.0$ is an ideal trained classifier, and $AUC=0.5$ does not
provide any classification information.  Our best performing models on
a training set of 10,000 lens containing images and 10,000 non-lens
containing images exhibit an AUC of 0.975 for an HST-like sample.
However, for one orbit of LSST, our model only reaches an AUC of
0.625.  For 10-year mock LSST observations, the AUC improves to 0.809.
Model performance appears to continually improve with the size of the
training set.  Models trained on fewer images perform better in
absence of the lens galaxy, which can decrease the false positive rate
by blending with light from a lensed galaxy image near the central
lensing galaxy.  However, with larger training data sets, information
from the lensing galaxy improves model performance.  Our
results demonstrate efficient and effective methods for automated strong
lensing identification. The simple classifier used here easily
scales to large data sets on a computing cluster and is easily
reproduced on a personal laptop, using existing open source tools
in both cases.
\end{abstract}
%-------------------------------------------------%

@@

%  Relevance of our work
While there has been a recent surge in the use of deep neural
networks applied to image classification problems in astronomy,
it is not always easy to scale these techniques to large data
sets, nor are the necessary computational resources and hardware,
such as GPUs, easily accessible to the
entire scientific community.  We present a supervised classification
pipeline built with open source tools, tested on mock HST and LSST
data, which will also be made publicly available.  Our goal is to
centralize an open source pipeline that is general enough for the
user to select or add appropriate image processing techniques to
augment the features of the lensed image, train and test data with
different machine learning techniques, and to select the sequence of
methods that provides optimal model performance.  Results are
reproducible on personal computers, as both the pipeline and the
data will be publicly distributed.

@@

% Benefits of our method
We emphasize that simple linear classifiers, such as
logistic regression, are scalable and relatively easy to parallelize
with open source tools such as {\em
Spark}\footnote{http://spark.apache.org/}.  Initial checks
indicate that HOG feature extraction plus a linear
classifier captures much of the complexity in the arc finding problem.
These results indicate that effective image classification of
galaxy-galaxy strong lensing systems can be achieved without
large scale computing resources, or specialized hardware like GPUs,
but can be scaled to large datasets on a computing cluster if needed.
